### 2016-03-01 abstract.tex
Initialize repo with the official template files

\begin{foreignabstract}

In this work, we present ...

\end{foreignabstract}