### 2016-03-01 paper1.tex
Add Latex converted text for fatiando paper
Fixed the figures and code samples. Still missing the references.

\chapter{Modeling the Earth with Fatiando a Terra}\label{chap:fatiando}

\section{Abstract}

Geophysics is the science of using physical observations of the Earth to
infer its inner structure. Generally, this is done with a variety of
numerical modeling techniques and inverse problems. The development of
new algorithms usually involves copy and pasting of code, which leads to
errors and poor code reuse. Fatiando a Terra is a Python library that
aims to automate common tasks and unify the modeling pipeline inside of
the Python language. This allows users to replace the traditional shell
scripting with more versatile and powerful Python scripting. The library
can also be used as an API for developing stand-alone programs.
Algorithms implemented in Fatiando a Terra can be combined to build upon
existing functionality. This flexibility facilitates prototyping of new
algorithms and quickly building interactive teaching exercises. In the
future, we plan to continuously implement sample problems to help teach
geophysics as well as classic and state-of-the-art algorithms.

\section{Introduction}

Geophysics studies the physical processes of the Earth. Geophysicists
make observations of physical phenomena and use them to infer the inner
structure of the planet. This task requires the numerical modeling of
physical processes. These numerical models can then be used in inverse
problems to infer inner Earth structure from observations. Different
geophysical methods use different kinds of observations. Geothermal
methods use the temperature and heat flux of the Earth's crust.
Potential field methods use gravitational and magnetic field
measurements. Seismics and seismology use the ground motion caused by
elastic waves from active (man-made) and passive (earthquakes) sources,
respectively.

The seismic method is among the most widely studied due to the high
industry demand. Thus, a range of well established open-source software
have been developed for seismic processing. These include
\href{http://www.cwp.mines.edu/cwpcodes/}{Seismic Un*x} (SU) {[}SU{]}\_,
\href{http://www.ahay.org/}{Madagascar} {[}MAD{]}\_,
\href{http://opendtect.org}{OpendTect}, and
\href{http://www.gebrproject.com}{GêBR}. A noteworthy open-source
project that is not seismic related is the
\href{http://gmt.soest.hawaii.edu/}{Generic Mapping Tools} (GMT) project
{[}GMT{]}\_. The GMT are a well established collection of command-line
programs for plotting maps with a variety of different map projections.
For geodynamic modeling there is the
\href{http://www.geodynamics.org}{Computational Infrastructure for
Geodynamics} (CIG), which has grouped various well documented software
packages. However, even with this wide range of well maintained software
projects, many geophysical modeling software that are provided online
still have no open-source license statement, have cryptic I/O files, are
hard to integrate into a pipeline, and make code reuse and remixing
challenging. Some of these problems are being worked on by the
\href{http://geosys.usc.edu/projects/seatree/}{Solid Earth Teaching and
Research Environment} (SEATREE) {[}SEATREE{]}\_ by providing a common
graphical interface for previously existing software. The numerical
computations are performed by the pre-existing underlying C/Fortran
programs. Conversely, the SEATREE code (written in Python) handles the
I/O and user interface. This makes the use of these tools easier and
more approachable to students. However, the lack of a common API means
that the code for these programs cannot be easily combined to create new
modeling tools.

\href{http://www.fatiando.org}{Fatiando a Terra} aims at providing such
an API for geophysical modeling. Functions in the \texttt{fatiando}
package use compatible data and mesh formats so that the output of one
modeling function can be used as input for another. Furthermore,
routines can be combined and reused to create new modeling algorithms.
Fatiando a Terra also automates common tasks such as griding, map
plotting with \href{http://matplotlib.org}{Matplotlib} {[}MPL{]}\_, and
3D plotting with
\href{http://code.enthought.com/projects/mayavi}{Mayavi} {[}MYV{]}\_.
Version 0.1 of Fatiando a Terra is focused on gravity and magnetic
methods because this is the main focus of the developers. However,
simple ``toy'' problems for seismology and geothermics are available and
can be useful for teaching geophysics.

The following sections illustrate the functionality and design of
Fatiando a Terra using various code samples. An
\href{http://ipython.org/}{IPython} {[}IPY{]}\_ notebook file with these
code samples is provided by {[}SAMPLES{]}\_ at
\url{http://dx.doi.org/10.6084/m9.figshare.708390}.

\section{Package structure}

The modules and packages of Fatiando a Terra are bundled into the
\texttt{fatiando} package. Each type of geophysical method has its own
package. As of version 0.1, the available modules and packages are:

\begin{itemize}
\item
  \texttt{fatiando.gravmag}: gravity and magnetic methods;
\item
  \texttt{fatiando.seismic}: seismic methods and seismology;
\item
  \texttt{fatiando.geothermal}: geothermal modeling;
\item
  \texttt{fatiando.mesher}: geometric elements and meshes;
\item
  \texttt{fatiando.gridder}: grid generation, slicing, interpolation,
  etc;
\item
  \texttt{fatiando.io}: I/O of models and data sets from web
  repositories;
\item
  \texttt{fatiando.utils}: miscellaneous utilities;
\item
  \texttt{fatiando.constants}: physical constants;
\item
  \texttt{fatiando.gui}: simple graphical user interfaces;
\item
  \texttt{fatiando.vis}: 2D and 3D plotting;
\item
  \texttt{fatiando.inversion}: inverse problem solvers and
  regularization;
\end{itemize}

\section{Griding and map plotting}

Fatiando a Terra handles map data as 1D Numpy arrays, typically x-, y-,
z-coordinates and an extra array with the corresponding data. However,
Matplotlib functions, like \texttt{contourf} and \texttt{pcolor},
require data to be passed as 2D arrays. Moreover, geophysical data sets
are often irregularly sampled and require griding before they can be
plotted. Thus, griding and array reshaping are ideal targets for
automation.

The \texttt{fatiando.vis.mpl} module imports all the functions in
\texttt{matplotlib.pyplot}, adds new functions, and overwrites others to
automate repetitive tasks (such as griding). Thus, the basic
functionality of the \texttt{pyplot} interface is maintained while
customizations facilitate common tasks. The following example
illustrates the use of the custom \texttt{fatiando.vis.mpl.contourf}
function to automatically grid and plot some irregularly sampled data
(Figure~\ref{fig:contourf}):

\begin{verbatim}
from fatiando import gridder
from fatiando.vis import mpl
area = [-20, 20, -50, 50]
x, y = gridder.scatter(area, n=100)
data = x**2 + y**2
mpl.figure()
mpl.axis('scaled')
mpl.contourf(y, x, data, shape=(50, 50),
    levels=30, interp=True)
mpl.colorbar(orientation='horizontal')
mpl.plot(y, x, '.k')
mpl.xlabel('y (East-West)')
mpl.ylabel('x (North-South)')
mpl.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/gridding_plotting_contourf}
    \caption{
    Example of 1) generating a random scatter of points (black dots),
    2) using that to make synthetic data, and
    3) automatically gridding and plotting the data using a
    Fatiando a Terra wrapper for the Matplotlib ``contourf``
    function.
    }
    \label{fig:contourf}
\end{figure}

Notice that, in the calls to \texttt{mpl.contourf} and
\texttt{mpl.plot}, the x- and y-axis are switched. That is because it is
common practice in geophysics for x to point North and y to point East.

Map projections in Matplotlib are handled by the
\href{http://matplotlib.org/basemap}{Basemap toolkit}. The
\texttt{fatiando.vis.mpl} module also provides helper functions to
automate the use of this toolkit. The \texttt{fatiando.vis.mpl.basemap}
function automates the creation of the \texttt{Basemap} objects with
common parameters. This object can then be passed to the
\texttt{contourf}, \texttt{contour} and \texttt{pcolor} functions in
\texttt{fatiando.vis.mpl} and they will automatically plot using the
given projection (Figure~\ref{fig:basemap}):

\begin{verbatim}
mpl.figure()
bm = mpl.basemap(area, projection='robin')
bm.drawmapboundary()
bm.drawcoastlines()
mpl.contourf(x, y, data, shape=(50, 50), levels=30,
    interp=True, basemap=bm)
mpl.colorbar(orientation='horizontal')
mpl.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/gridding_plotting_basemap}
    \caption{
        Example of map plotting with the Robinson projection using the
        Matplotlib Basemap toolkit.
    }
    \label{fig:basemap}
\end{figure}

\section{Meshes and 3D plotting}

The representation of 2D and 3D geometric elements is handled by the
classes in the \texttt{fatiando.mesher} module. Geometric elements in
Fatiando a Terra can be assigned physical property values, like density,
magnetization, seismic wave velocity, impedance, etc. This is done
through a \texttt{props} dictionary whose keys are the name of the
physical property and values are the corresponding values in SI units:

\begin{verbatim}
from fatiando import mesher
model = [
    mesher.Prism(5, 8, 3, 7, 1, 7,
        props={'density':200}),
    mesher.Prism(1, 2, 4, 5, 1, 2,
        props={'density':1000})]
\end{verbatim}

The \texttt{fatiando.vis.myv} module contains functions to automate 3D plotting
using Mayavi {[}MYV{]}\_. The \texttt{mayavi.mlab} interface requires geometric
elements to be formatted as TVTK objects. Thus, plotting functions in
\texttt{fatiando.vis.myv} automatically create TVTK representations of
\texttt{fatiando.mesher} objects and plot them using a suitable function of
\texttt{mayavi.mlab}. Also included are utility functions for drawing axes,
walls on the figure bounding box, etc. For example, the
\texttt{fatiando.vis.myv.figure} function creates a figure and rotates it so
that the z-axis points down, as is standard in geophysics. The following
example shows how to plot the 3D right rectangular prism model that we created
previously (Figure~\ref{fig:twoprisms}):

\begin{verbatim}
from fatiando.vis import myv
bounds = [0, 10, 0, 10, 0, 10]
myv.figure()
myv.prisms(model, 'density')
myv.axes(myv.outline(bounds))
myv.wall_bottom(bounds)
myv.wall_north(bounds)
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_2prisms}
    \caption{
        Example of plotting a list of right rectangular prisms in Mayavi.
    }
    \label{fig:twoprisms}
\end{figure}

The \texttt{fatiando.mesher} module also contains classes for
collections of elements (e.g., meshes). A good example is the
\texttt{PrismMesh} class that represents a structured mesh of right
rectangular prisms. This class behaves as a list of
\texttt{fatiando.mesher.Prism} objects and can be passed to functions
that ask for a list of prisms, like \texttt{fatiando.vis.myv.prisms}.
Physical properties can be assigned to the mesh using the
\texttt{addprop} method (Figure~\ref{fig:mesh}):

\begin{verbatim}
mesh = mesher.PrismMesh(bounds, shape=(3, 3, 3))
mesh.addprop('density', range(mesh.size))
myv.figure()
myv.prisms(mesh, 'density')
myv.axes(myv.outline(bounds))
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_mesh}
    \caption{
        Example of generating and visualizing a structured prism mesh.
    }
    \label{fig:mesh}
\end{figure}

Often times the mesh is used to make a detailed model of an irregular
region of the Earth's surface. In such cases, it is necessary to
consider the topography of the region. The \texttt{PrismMesh} class has
a \texttt{carvetopo} method that masks the prisms that fall above the
topography. The example below illustrates this functionality using
synthetic topography (Figure~\ref{fig:meshtopo}):

\begin{verbatim}
from fatiando import utils
x, y = gridder.regular(bounds[:4], (50, 50))
heights = -5 + 5*utils.gaussian2d(x, y, 10, 5,
    x0=10, y0=10)
mesh = mesher.PrismMesh(bounds, (20, 20, 20))
mesh.addprop('density', range(mesh.size))
mesh.carvetopo(x, y, heights)
myv.figure()
myv.prisms(mesh, 'density')
myv.axes(myv.outline(bounds))
myv.wall_north(bounds)
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_meshtopo}
    \caption{
        Example of generating and visualizing a prism mesh with masked
        topography.
    }
    \label{fig:meshtopo}
\end{figure}

When modeling involves the whole Earth, or a large area of it, the
geophysicist needs to take into account the Earth's curvature. In such
cases, rectangular prisms are inadequate for modeling and tesseroids
(e.g., spherical prisms) are better suited. The
\texttt{fatiando.vis.myv} module contains auxiliary functions to plot
along with tesseroids: an Earth-sized sphere, meridians and parallels,
as well as continental borders (Figure~\ref{fig:tesseroid}):

\begin{verbatim}
model = [
    mesher.Tesseroid(-60, -55, -30, -27, 500000, 0,
        props={'density':200}),
    mesher.Tesseroid(-66, -55, -20, -10, 300000, 0,
        props={'density':-100})]
fig = myv.figure(zdown=False)
myv.tesseroids(model, 'density')
myv.continents(linewidth=2)
myv.earth(opacity=1)
myv.meridians(range(0, 360, 45), opacity=0.2)
myv.parallels(range(-90, 90, 45), opacity=0.2)
# Rotate the camera to get a good view
scene = fig.scene
scene.camera.position = [21199620.406122234,
    -12390254.839673528, -14693312.866768979]
scene.camera.focal_point = [-535799.97230670298,
    -774902.33205294283, 826712.82283183688]
scene.camera.view_angle = 19.199999999999996
scene.camera.view_up = [0.33256519487680014,
    -0.47008782429014295, 0.81756824095039038]
scene.camera.clipping_range = [7009580.0037488714,
    55829873.658824757]
scene.camera.compute_view_plane_normal()
scene.render()
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_tesseroid}
    \caption{
        Example of creating a tesseroid (spherical prism) model and visualizing
        it in Mayavi.
    }
    \label{fig:tesseroid}
\end{figure}

\section{Forward modeling}

In geophysics, the term ``forward modeling'' is used to describe the
process of generating synthetic data from a given Earth model.
Conversely, geophysical inversion is the process of estimating Earth
model parameters from observed data.

The Fatiando a Terra packages have separate modules for forward modeling
and inversion algorithms. The forward modeling functions usually take as
arguments geometric elements from \texttt{fatiando.mesher} with assigned
physical properties and return the synthetic data. For example, the
module \texttt{fatiando.gravmag.tesseroid} is a Python implementation of
the program Tesseroids (\url{http://leouieda.github.io/tesseroids}) and
calculates the gravitational fields of tesseroids (e.g., spherical
prisms). The following example shows how to calculate the gravity
anomaly of the tesseroid model generated in the previous section
(Figure~\ref{fig:tesseroidgrav}):

\begin{verbatim}
from fatiando import gravmag
area = [-80, -30, -40, 10]
shape = (50, 50)
lons, lats, heights = gridder.regular(area, shape,
    z=2500000)
gz = gravmag.tesseroid.gz(lons, lats, heights, model)
mpl.figure()
bm = mpl.basemap(area, 'ortho')
bm.drawcoastlines()
bm.drawmapboundary()
bm.bluemarble()
mpl.title('Gravity anomaly (mGal)')
mpl.contourf(lons, lats, gz, shape, 30, basemap=bm)
mpl.colorbar()
mpl.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_tesseroid_data}
    \caption{
        Example of forward modeling the gravity anomaly using the tesseroid
        model shown in Figure~\ref{fig:tesseroid}.
    }
    \label{fig:tesseroidgrav}
\end{figure}

The module \texttt{fatiando.gravmag.polyprism} implements the method of
{[}PLOUFF{]}\_ to forward model the gravity fields of a 3D right
polygonal prism. The following code sample shows how to interactively
generate a polygonal prism model and calculate its gravity anomaly
(Figures~\ref{fig:drawing} and \ref{fig:polyprism}):

\begin{verbatim}
# Draw a polygon and make a polygonal prism
bounds = [-1000, 1000, -1000, 1000, 0, 1000]
area = bounds[:4]
mpl.figure()
mpl.axis('scaled')
vertices = mpl.draw_polygon(area, mpl.gca(),
    xy2ne=True)
model = [mesher.PolygonalPrism(vertices, z1=0,
    z2=500, props={'density':500})]
# Calculate the gravity anomaly
shape = (100, 100)
x, y, z = gridder.scatter(area, 300, z=-1)
gz = gravmag.polyprism.gz(x, y, z, model)
mpl.figure()
mpl.axis('scaled')
mpl.title("Gravity anomaly (mGal)")
mpl.contourf(y, x, gz, shape=(50, 50),
    levels=30, interp=True)
mpl.colorbar()
mpl.polygon(model[0], '.-k', xy2ne=True)
mpl.set_area(area)
mpl.m2km()
mpl.show()
myv.figure()
myv.polyprisms(model, 'density')
myv.axes(myv.outline(bounds),
        ranges=[i*0.001 for i in bounds])
myv.wall_north(bounds)
myv.wall_bottom(bounds)
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/forward_modeling_polyprism_drawing}
    \caption{
        Screen-shot of interactively drawing the contour of a 3D polygonal
        prism, as viewed from above.
    }
    \label{fig:drawing}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/paper1/forward_modeling_polyprism}
    \caption{
        Example of forward modeling the gravity anomaly of a 3D polygonal
        prism.
        a) forward modeled gravity anomaly.
        b) 3D plot of the polygonal prism.
    }
    \label{fig:polyprism}
\end{figure}

\section{Gravity and magnetic methods}

Geophysics uses anomalies in the gravitational and magnetic fields
generated by density and magnetization contrasts within the Earth to
investigate the inner Earth structure. The Fatiando a Terra 0.1 release
has been focused on gravity and magnetic methods. Therefore, the
\texttt{fatiando.gravmag} package contains more advanced and
state-of-the-art algorithms than the other packages.

The module \texttt{fatiando.gravmag.imaging} implements the imaging
methods described in {[}FP{]}\_. These methods aim to produce an image
of the geologic source from the observed gravity or magnetic data. The
following code sample uses the ``sandwich model'' method {[}SNDW{]}\_ to
image the polygonal prism, produced in the previous section, based on
its gravity anomaly (Figure~\ref{fig:imaging}):

\begin{verbatim}
estimate = gravmag.imaging.sandwich(x, y, z, gz,
    shape, zmin=0, zmax=1000, nlayers=20, power=0.2)
body = mesher.vfilter(1.3*10**8, 1.7*10**8,
    'density', estimate)
myv.figure()
myv.prisms(body, 'density', edges=False)
p = myv.polyprisms(model, 'density',
    style='wireframe', linewidth=4)
p.actor.mapper.scalar_visibility = False
p.actor.property.color = (0, 0, 0)
myv.axes(myv.outline(bounds),
    ranges=[i*0.001 for i in bounds])
myv.wall_north(bounds)
myv.wall_bottom(bounds)
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_imaging}
    \caption{
        Example of using the "sandwich model" imaging method to recover a 3D
        image of a geologic body based on its gravity anomaly. The colored
        blocks are a cutoff of the imaged body. The black contours are the true
        source of the gravity anomaly.
    }
    \label{fig:imaging}
\end{figure}

Also implemented in Fatiando a Terra are some recent developments in
gravity and magnetic inversion methods. The method of ``planting
anomalous densities'' by {[}UB{]}\_ is implemented in the
\texttt{fatiando.gravmag.harvester} module. In contrast to imaging
methods, this is an inversion method, i.e., it estimates a physical
property distribution (density in the case of gravity data) that fits
the observed data. This particular method requires the user to specify a
``seed'' (Figure~\ref{fig:seed}) around which the estimated density
distribution grows (Figure~\ref{fig:harvester}):

\begin{verbatim}
# Make a mesh and a seed
mesh = mesher.PrismMesh(bounds, (15, 30, 30))
seeds = gravmag.harvester.sow(
    [[200, 300, 100, {'density':500}]],
    mesh)
myv.figure()
myv.prisms([mesh[s.i] for s in seeds])
p = myv.polyprisms(model, 'density',
    style='wireframe', linewidth=4)
p.actor.mapper.scalar_visibility = False
p.actor.property.color = (0, 0, 0)
myv.axes(myv.outline(bounds),
    ranges=[i*0.001 for i in bounds])
myv.wall_north(bounds)
myv.wall_bottom(bounds)
myv.show()
# Now perform the inversion
data = [gravmag.harvester.Gz(x, y, z, gz)]
estimate = gravmag.harvester.harvest(data, seeds,
    mesh, compactness=0.1, threshold=0.0001)[0]
mesh.addprop('density', estimate['density'])
body = mesher.vremove(0, 'density', mesh)
myv.figure()
myv.prisms(body, 'density')
p = myv.polyprisms(model, 'density',
    style='wireframe', linewidth=4)
p.actor.mapper.scalar_visibility = False
p.actor.property.color = (0, 0, 0)
myv.axes(myv.outline(bounds),
    ranges=[i*0.001 for i in bounds])
myv.wall_north(bounds)
myv.wall_bottom(bounds)
myv.show()
\end{verbatim}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_harvester_seed}
    \caption{
        The small blue prism is the seed used by
        \texttt{fatiando.gravmag.harvester} to perform the inversion of a
        gravity anomaly. The black contours are the true source of the gravity
        anomaly.
    }
    \label{fig:seed}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_harvester}
    \caption{
        The blue prisms are the result of a gravity inversion using module
        \texttt{fatiando.gravmag.harvester}. The black contours are the true
        source of the gravity anomaly. Notice how the inversion was able to
        recover the approximate geometry of the true source.
    }
    \label{fig:harvester}
\end{figure}

\section{A toy seismic tomography}

The following example uses module \texttt{fatiando.seismic.srtomo} to
perform a simplified 2D tomography on synthetic seismic wave travel-time
data. To generate the travel-times we used a seismic wave velocity model
constructed from an image file. The colors of the image are converted to
gray-scale and the intensity is mapped to seismic wave velocity by the
\texttt{img2prop} method of the \texttt{fatiando.mesher.SquareMesh}
class. This model (Figure~\ref{fig:tomo}) is then used to calculate the
travel-times between a random set of earthquake locations and seismic
receivers (seismometers):

\begin{verbatim}
import urllib
from fatiando import mesher, utils, seismic
from fatiando.vis import mpl
area = (0, 500000, 0, 500000)
shape = (30, 30)
model = mesher.SquareMesh(area, shape)
link = '/'.join(["http://fatiando.readthedocs.org",
    "en/Version0.1/_static/logo.png"])
urllib.urlretrieve(link, 'model.png')
model.img2prop('model.png', 4000, 10000, 'vp')
quake_locations = utils.random_points(area, 40)
receiver_locations = utils.circular_points(area, 20,
    random=True)
quakes, receivers = utils.connect_points(
    quake_locations, receiver_locations)
traveltimes = seismic.ttime2d.straight(model, 'vp',
    quakes, receivers)
noisy = utils.contaminate(traveltimes, 0.001,
    percent=True)
\end{verbatim}

Now the noise-corrupted synthetic travel-times can be used in our
simplified tomography:

\begin{verbatim}
mesh = mesher.SquareMesh(area, shape)
slowness, residuals = seismic.srtomo.run(noisy,
    quakes, receivers, mesh, smooth=10**6)
velocity = seismic.srtomo.slowness2vel(slowness)
mesh.addprop('vp', velocity)
# Make the plots
mpl.figure(figsize=(9, 7))
mpl.subplots_adjust(top=0.95, bottom=0.05,
    left=0.05, right=0.95)
mpl.subplot(2, 2, 1)
mpl.title('Velocity model (m/s)')
mpl.axis('scaled')
mpl.squaremesh(model, prop='vp', cmap=mpl.cm.seismic)
mpl.colorbar(pad=0.01)
mpl.points(quakes, '*y', label="Sources")
mpl.points(receivers, '^g', label="Receivers")
mpl.m2km()
mpl.subplot(2, 2, 2)
mpl.title('Ray paths')
mpl.axis('scaled')
mpl.squaremesh(model, prop='vp', cmap=mpl.cm.seismic)
mpl.colorbar(pad=0.01)
mpl.paths(quakes, receivers)
mpl.points(quakes, '*y', label="Sources")
mpl.points(receivers, '^g', label="Receivers")
mpl.m2km()
mpl.subplot(2, 2, 3)
mpl.title('Estimated velocity (m/s)')
mpl.axis('scaled')
mpl.squaremesh(mesh, prop='vp', cmap=mpl.cm.seismic,
    vmin=4000, vmax=10000)
mpl.colorbar(pad=0.01)
mpl.m2km()
mpl.subplot(2, 2, 4)
mpl.title('Residuals (s)')
mpl.hist(residuals, bins=10)
mpl.show()
\end{verbatim}

Even though the implementation in \texttt{fatiando.seismic.srtomo} is greatly
simplified and not usable in real tomography problems, the result in
Figure~\ref{fig:tomo} illustrates interesting inverse problem concepts.  Notice
how the estimated velocity is blurred in the corners where no rays pass
through. This is because the data (travel-times) provide no information about
the velocity in those areas. Areas like those constitute the null space of the
inverse problem {[}MENKE{]}\_, where any velocity value estimated will provide
an equal fit to the data.  Thus, the tomography problem requires the use of
prior information in the form of regularization. Most commonly used in
tomography problems is the Tikhonov first-order regularization, e.g., a
smoothness constraint {[}MENKE{]}\_. The amount of smoothness imposed on the
solution is controlled by the \texttt{smooth} argument of function
\texttt{fatiando.seismic.srtomo.run}. That is how we are able to estimate a
unique and stable solution and why the result is specially smoothed where there
are no rays.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/seismic_tomo}
    \caption{
        Example run of a simplified 2D tomography. The top-left panel shows the
        true velocity model with the locations of earthquakes (yellow stars)
        and receivers (green triangles). The top-right panel shows the
        ray-paths between earthquakes and receivers. The bottom-left panel is
        the velocity estimated by the tomography. The bottom-right panel is a
        histogram of the travel-time residuals of the tomography. Notice how
        the majority of residuals are close to 0 s, indicating a good fit to
        the data.
    }
    \label{fig:tomo}
\end{figure}

\section{Conclusion}

The Fatiando a Terra package provides an API to develop modeling
algorithms for a variety of geophysical methods. The current version
(0.1) has a few state-of-the-art gravity and magnetic modeling and
inversion algorithms. There are also toy problems in gravity, seismics
and seismology that are useful for teaching basic concepts of
geophysics, modeling, and inverse problems.

Fatiando a Terra enables quick prototyping of new algorithms because of
the collection of fast forward modeling routines and the simple syntax
and high level of the Python language. After prototyping, the
performance bottlenecks of these algorithms can be easily diagnosed
using the advanced profiling tools available in the Python language.
Optimization of only small components of code can be done without loss
of flexibility using the Cython language {[}CYTHON{]}\_.

The biggest challenge that Fatiando a Terra faces in the near future is
the development of a user and, consequently, a developer community. This
is a key part for the survival of any open-source project.

\section{Acknowledgments}

The authors were supported by a scholarship (L. Uieda) from Coordenação de
Aperfeiçoamento de Pessoal de Nível Superior (CAPES), a scholarship (V.C.
Oliveira Jr) from Conselho Nacional de Desenvolvimento Científico e Tecnológico
(CNPq), and a fellowship (V.C.F. Barbosa) from CNPq.  Additional support was
provided by the Brazilian agencies CNPq (grant 471693/2011-1) and FAPERJ (grant
E-26/103.175/2011).