### 2016-03-17 conclusion.tex
Cite papers that used fatiando

\chapter{Conclusions}

We have developed two open-source software packages that implement the
components required for building an inversion method: forward modeling,
optimization, and regularization.
We used these components to develop a fast gravity inversion method to estimate
the depth of the crust-mantle interface (the Moho) in a spherical
approximation.
We then applied the proposed method to estimate the Moho depth for South
America.

The \textit{Tesseroids} software is a collection of command-line
programs developed in the C programming language.
The programs calculate the gravitational potential and its first and second
derivatives for a tesseroid (spherical prism) model.
We implemented an adaptive discretization algorithm to guarantee the accuracy
of the computations.
The adaptive discretization is controlled by the distance-size
ratio ($D$).
Higher values of $D$ result in finner discretization and vice-versa.
Furthermore, we investigated the accuracy of the calculations as a function of
$D$.
Contrary to previous assumptions, our results showed that the first and second
derivatives require finner discretization than the gravitational potential to
achieve the same accuracy level.
The values of the distance-size ratio that yield a maximum error of 0.1\%
are $D = 1$ for the gravitational potential, $D = 1.5$ for the first
derivatives, and $D = 8$ for the second derivatives.

\textit{Fatiando a Terra} is a library implemented in the Python programming
language.
The library contains functions and classes for data processing, visualization,
forward modeling, and inversion.
A new inversion method can be implemented by combining the existing forward
modeling functions with the problem agnostic optimization and regularization
tools in the library.
Using these tools, the amount of code required to implement new methods is
reduced, increasing the speed of the cycle of prototyping a new algorithm,
testing, and then refining it.
The project has been used in scholarly works such as
\citet{hidalgo-gato2015a}, \citet{niccoli2015}, \citet{oliveirajr.2015},
and \citet{bassett2016}.
To date, the project has received contributions from nine developers in three
different countries.